### 2022-05-31 paper.Rmd
work on abstract

    Anticlustering refers to a collection of methods that partition data sets into disjoint groups while aiming for high between-group similarity and high within-group heterogeneity. Anticlustering is usually approached by maximzing instead of minimizing a clustering objective such as the k-means criterion. Introducing k-plus, this paper presents an extension of k-means to maximize between-group similarity. While k-means anticlustering minimizes between-group-differences only with regard to the mean of attribute values, k-plus anticlustering can be used to minimize between-group-differences with regard to the variance, higher order moments like skewness and kurtosis, and even covariance structure. It is shown that all k-plus variations can be reduced to an augmentation of the input data while the original k-means objective remains unchanged. A computer simulation and practical examples on real data show that k-plus anticlustering achieves high similarity with regard to mutliple distribution characteristics with little sacrifice differences in means. The k-plus extension is therefore preferred over classical k-means anticlustering when optimizing between-group similarity. Example are given on how k-plus anticlustering can be applied using the open source R package `anticlust`, which is freely available via CRAN (https://cran.r-project.org/package=anticlust) and Github (https://github.com/m-Py/anticlust).

@@

Anticlustering is a collection of methods to partition a set of elements into disjunct groups, with the goal of obtaining high between-group similarity or within-group heterogeneity [@brusco2019; @spath1986; @valev1998]. Oftentimes, both goals coincide: some anticlustering objectives imply that similarity between groups is maximal whenever similarity within groups is minimal [@brusco2019; @papenberg2020; @feo1990]. Anticlustering thereby reverses the logic of its better known twin---cluster analysis---which seeks homogeneity within clusters and heterogeneity between clusters [@rokach2005; @steinley2006]. Whereas a variety of different approaches exist for cluster analysis such as hierarchical methods, model-based clustering, or network approaches [see @brusco_emergent_2012, for an overview], anticlustering is usually formalized as a partitioning process that divides a set of elements into disjunct groups in such a way that an objective function---representing between-group similarity and/or within-group heterogeneity---is optimized [@papenberg2020; @spath1986; @brusco2019].

@@

## Closing thoughts on the nature of anticlustering

- new criterion: optimizes between-group similarity, no correspondence to within-group heterogeneity [see @brusco2019 for a exceptional method maximizing within-group heterogeneity]

- Similarly: the reversed objective no longer has a useful clustering interpretation: Maximizing differences in variances does not lead to homogeneous and/or well separated clusters

>>>>

    Anticlustering refers to a collection of methods that partition data sets into disjoint groups while aiming for high between-group similarity or high within-group heterogeneity. Anticlustering is usually approached by maximzing instead of minimizing a clustering objective such as the sum of squared errors used in k-means clustering. Introducing k-plus, this paper presents an extension of the k-means criterion to maximize between-group similarity. While k-means anticlustering minimizes between-group-differences only with regard to the mean of attribute values, k-plus anticlustering can be used to minimize differences with regard to multiple distribution characteristics: the variance, higher order moments such as skewness and kurtosis, and even to preserve covariance structure. Thus, k-plus corresponds to a family of new anticlustering criteria that can be combined in arbitrary ways. It is shown that all k-plus variations can be reduced to an augmentation of the input data while the original k-means objective remains unchanged. A computer simulation and practical examples show that k-plus anticlustering achieves high between-group similarity with regard to multiple criteria with little sacrifice to each single criterion. The k-plus extension is therefore preferred over classical k-means anticlustering when optimizing between-group similarity. Example are given on how k-plus anticlustering can be applied to real norming data using the open source R package `anticlust`, which is freely available via CRAN (https://cran.r-project.org/package=anticlust) and Github (https://github.com/m-Py/anticlust).

@@

Anticlustering is a collection of methods to partition a set of elements into disjoint groups, with the goal of obtaining high between-group similarity or within-group heterogeneity [@brusco2019; @spath1986; @valev1998]. Oftentimes, both goals coincide: some anticlustering objectives imply that similarity between groups is maximal whenever similarity within groups is minimal [@brusco2019; @papenberg2020; @feo1990]. Anticlustering thereby reverses the logic of its better known twin---cluster analysis---which seeks homogeneity within clusters and heterogeneity between clusters [@rokach2005; @steinley2006]. Whereas a variety of different approaches exist for cluster analysis such as hierarchical methods, model-based clustering, or network approaches [see @brusco_emergent_2012, for an overview], anticlustering is usually formalized as a partitioning process that divides a set of elements into disjoint groups in such a way that an objective function---representing between-group similarity and/or within-group heterogeneity---is optimized [@papenberg2020; @spath1986; @brusco2019].

@@
